Relict deciduous oaks show high resilience to drought despite the effect of land-use legacies

### 2020-05-24 analysis/evi/compute_anomalies.Rmd
update

---
title: "Compute EVI anomalies"
author: "AJ Perez-Luque (@ajpelu)"
date: "2017 Aug"
output:  
  md_document:
    variant: markdown_github
---
```{r, echo=FALSE, message=FALSE}
require(knitr)
opts_chunk$set(message = FALSE, warning = FALSE) 
```

@@

@@

write.csv(anomalias_evimean, file=paste(di, "/data/anomalies/anomalias_evimean.csv", sep=""), row.names = FALSE)
```

@@

write.csv(anomalias_composite, file=paste(di, "/data/anomalies/anomalias_composite.csv", sep=""), row.names = FALSE)
```

>>>>

---
title: "Compute EVI anomalies"
author: "AJ Perez-Luque (@ajpelu)"
date: "2017 Aug"
output:  
  md_document:
    variant: markdown_github
editor_options: 
  chunk_output_type: console
---
```{r, echo=FALSE, message=FALSE}
require(knitr)
opts_chunk$set(message = FALSE, warning = FALSE) 
```

@@

    for (y in years){
      # get mean evi for the reference period
      iv_ref <- df_aux %>% 
        # filter(year != y) %>% 
        summarise(mean(evi, na.rm=T))

      # Get sd ref for standardized anomaly
      sd_ref <- df_aux %>% 
        # filter(year != y) %>% 
        summarise(sd(evi, na.rm=T))

@@

write.csv(anomalias_evimean, file=paste(di, "/data/anomalies/anomalias_evimean_corrected.csv", sep=""), row.names = FALSE)

# write.csv(anomalias_evimean, file=paste(di, "/data/anomalies/anomalias_evimean.csv", sep=""), row.names = FALSE)

```

@@

write.csv(anomalias_composite, file=paste(di, "/data/anomalies/anomalias_composite.csv", sep=""), row.names = FALSE)

a <- avg_sa_clu %>% 
  mutate(method = "original") %>% 
  bind_rows(avg_sa_cluc %>% mutate(method = "no remove"))

```

```{r, eval=FALSE, echo=FALSE}
df_aux <- evi[evi$iv_malla_modi_id == 144160, ]

years <- unique(df$year)

      # Create dataframe 
      aux <- cbind(y, iv_year, iv_ref, anomaly, anomaly_nor, anomaly_std)
      aux_anomaly <- rbind(aux_anomaly, aux)
      }

  # Add name of the pixel 
  aux_anomaly <- aux_anomaly %>% 
    mutate(iv_malla_modi_id = i) %>% 
    mutate(evi_mean_cte = mean(iv_year))

```

### 2020-05-24 man/figures/0_figures_ecosystems.Rmd
update

@@

![](../../man/figures/svg/fig_7.jpg)

**Figure 7.** Correlation coefficients found by relating tree-ring residual chronologies (RWI) of *Q. pyrenaica* and monthly climatic data: precipitation and 6-month SPEI (a), minimum (b) and maximum (c) temperatures. *green* bars: northern site (SJ); *light blue* bars: low-elevation southern site (CA-Low); and *dark blue* bars: high-elevation southern site (CA-High). Asterisks indicate significant ($P < 0.05$) correlation coefficients.

![](../../man/figures/svg/fig_8.jpg)

![](../../man/figures/svg/fig_s1ab.jpg)

![](../../man/figures/svg/fig_s2ab.jpg)

![](../../man/figures/svg/fig_s4ab.jpg)

##### 
**Figure S4.** EVI annual profile (average of the period 2000-2016) for *Q. pyrenaica* forests in Sierra Nevada and drought events. Horizontal bars correspond to the most severe droughts for Sierra Nevada since 1900 (computed as in Table S3). Their position indicates the start and end months of each drought event. Bars lengths show the duration of the drought event (number of consecutive months with SPEI lower than -1.28, see [@Pascoa2017].

**Figure S5.** Resilience metrics of the tree growth for severe drought events since 1950 (excluding 1995 drought event). *Left*: Resistance (*Rt*); *Center*: Recovery (*Rc*); *Right*: Resilience (*Rs*). Points indicate resilience metrics for oak populations: SJ (*blue*), CA-High (*red*) and CA-Low (*green*). Resilience metrics were computed for each population (sample depth > 10) and drought event. The gray line represents overall relationship for each Resilience metrics.

![](../../man/figures/svg/fig_s8.jpg)

>>>>

**Figure 3.** ***a)*** Basal Area Increment (BAI) chronologies of *Q. pyrenaica* for northern population (SJ; *red*) and southern ones: low-elevation (CA-Low; *green*) and high-elevation (CA-High, *blue*) sites. Shading areas correspond to standard error of the mean. Number of series is displayed in the upper plot. Only years replicated with # series > 5 are shown. Linear trends since 1975 are indicated for all sites (numbers indicate $r^2$ values; asterisks indicate significant linear trend, *p < 0.001*). ***b)*** Comparison of median growth change ($GC$) following @Nowacki1997 for *Q. pyrenaica* sites. Dashed black lines indicate a threshold of 50% of GC (see Material and methods). Note that y-axes do not correspond in all of the three panels for the sake of clarity. Error bars indicate standard error.

**Figure 4.** ***a)*** Resilience metrics of tree-growth for eight severe drought events since 1950 (see main text for details) as a function of drought severity. Points indicate resilience metrics for oak populations: SJ (*red*), CA-High (*blue*) and CA-Low (*green*). Resilience metrics were computed for each population (sample depth > 10) and drought event. Gray lines represent overall relationships for each Resilience metrics. ***b)*** Comparison of the response of  *Q. pyrenaica* forests to drought in terms of resistance, recovery, and resilience of greenness (b) and tree growth (c). For EVI, northern populations (*red circle*) were compared with southern ones (*green circle*). For BAI, the more xeric northern population (San Juan, SJ; *red circle*) was compared with the two southern populations, Cáñar-High (CA-High; *blue circle*) and Cáñar-Low (CA-Low; *green circle*). Different letters indicate significant *post hoc* differences between groups (see Material and methods for details).

#####
**Figure 1**
![](../../out/fig/geomap/mapa_situacion_v2.jpg){height=750px} 
#####
**Figure 2.**
![](../../man/figures/svg/fig_2ab.jpg)

#####
**Figure 3**

![](../../man/figures/svg/cronos_bai_gc.jpg)

#####
**Figure 4**
![](../../man/figures/svg/resiliences.jpg)

**Figure S4.** EVI annual profile (average of the period 2000-2016) for *Q. pyrenaica* forests in Sierra Nevada and drought events. Horizontal bars correspond to the most severe droughts for Sierra Nevada since 1900 (computed as in Table S3). Their position indicates the start and end months of each drought event. Bars lengths show the duration of the drought event (number of consecutive months with SPEI lower than -1.28, see [@Pascoa2017].

**Figure S5.** Resilience metrics of the tree growth for severe drought events since 1950 (excluding 1995 drought event). *Left*: Resistance; *Center*: Recovery; *Right*: Resilience. Points indicate resilience metrics for oak populations: SJ (*blue*), CA-High (*red*) and CA-Low (*green*). Resilience metrics were computed for each population (sample depth > 10) and drought event. The gray line represents overall relationship for each Resilience metrics.

**Figure S6.** Correlation coefficients found by relating tree-ring residual chronologies (RWI) of *Q. pyrenaica* and monthly climatic data: precipitation and 6-month SPEI (a), minimum (b) and maximum (c) temperatures. *green* bars: northern site (SJ); *light blue* bars: low-elevation southern site (CA-Low); and *dark blue* bars: high-elevation southern site (CA-High). Asterisks indicate significant ($P < 0.05$) correlation coefficients.

**Figure S1**
![](../../man/figures/svg/fig_s1ab.jpg)

##### 
**Figure S2**
![](../../man/figures/svg/fig_s2ab.jpg)

##### 
**Figure S3**
![](../../man/figures/svg/fig_s4ab.jpg)

##### 
**Figure S4**

##### 
**Figure S5**
![](../../man/figures/svg/fig_s8.jpg)

#####
**Figure S6**
![](../../man/figures/svg/fig_8.jpg)

```{r, eval=FALSE, echo=FALSE}

OLD

@@

![](../../man/figures/svg/fig_7.jpg)
```

### 2020-05-24 man/figures/std_anomalies.Rmd
update

>>>>

```{r}
## Reviewer 1. Is it necessary/correct to remove year i from the calculation of EVImean, ref?

# From explore_anomalies.Rmd 
# Read data 
anomalias_evimeanc <- read.csv(here("data/anomalies", "anomalias_evimean_corrected.csv"),header = TRUE, sep = ',')

anomalias_evimeanc <- anomalias_evimeanc %>% 
  mutate(
    clu_pop = as.factor(case_when(
      pop == 1 ~ "Camarate",
      pop %in% c(2,3,4,5) ~ 'Northern slope',
      pop %in% c(6,7,8) ~ 'Southern slope',
      pop == 9 ~ 'out')),
    clu_pop2 = as.factor(case_when(
      pop %in% c(1,2,3,4,5) ~ 'Northern slope',
      pop %in% c(6,7,8) ~ 'Southern slope',
      pop == 9 ~ 'out'))) %>% 
  filter(clu_pop != 'out')

# Standardized Anomalies 
avg_sa_cluc <- anomalias_evimeanc %>% 
  group_by(clu_pop2, y) %>% 
  dplyr::summarise(mean=mean(sa, na.rm=T),
            sd = sd(sa, na.rm=T),
            se = sd/sqrt(length(sa))) %>% 
 mutate(signo = ifelse(mean >= 0, 'pos', 'neg'))

a <- avg_sa_clu %>% 
  mutate(method = "original") %>% 
  bind_rows(avg_sa_cluc %>% mutate(method = "no remove")) %>% 
  unite("f", signo:method, remove = FALSE)

comparative_anomalies <- a %>% ggplot(aes(x=y, y=mean, fill=f)) + 
  geom_hline(yintercept = 0, size=0.3, colour = "gray") + 
  geom_bar(position = "dodge", stat = "identity", width = 0.8) + geom_errorbar(aes(ymin = mean - se, ymax = mean + se, colour = f), 
                width=.5, position = position_dodge(0.9)) +
  facet_wrap(~clu_pop2, nrow = 2) + 
  scale_fill_manual(values = c("neg_original" =  "#a63603",
                               "neg_no remove" =  "#FFCCCC",
                               "pos_original" = "#006d2c", 
                               "pos_no remove" = "#CCFFCC"
                                 )) +
  scale_colour_manual(values = c("neg_original" =  "#a63603",
                               "neg_no remove" =  "#FFCCCC",
                               "pos_original" = "#006d2c", 
                               "pos_no remove" = "#CCFFCC"
                                 )) +
  scale_x_continuous(breaks = seq(2000, 2016, by=2)) +
  ylab('EVI Standardized Anomaly') + xlab('Year') +
  theme_bw() +
  theme(axis.text = element_text(size=8),
        legend.position = "none",
        panel.grid = element_blank(),
        strip.background = element_rect(fill = "transparent", colour = "black", size = 0.2),
        plot.title = element_text(hjust = -0.05, size = 9, 
                                    margin = margin(t=-5, b=1)))
```
